Quanta concentration and Transmission risk

### 2024-03-04 analysis/qctr.Rmd
QCTR results draft

source("../utils/plotting.r")
source("../utils/tex.r")
source("../models/stm-v2.R")

# building
cellSize = 250
source("../preprocessing/prep_building-rasterize.R")
```

@@

```{r}
# data
mth = 10
dy = 25
sim = 1
param_1025 <- readRDS(paste0("../simulations/", "/", "sim-param-", 2021, "-", mth, "-", dy, ".rds"))
track_1025 <- readRDS(paste0("../simulations/", "/", "tracking-df-", 250, "px-", 2021, "-", mth, "-", dy, ".rds"))
track_undiag <- track_1025 %>%
  dplyr::filter(tracking_end != "Possible HCW") # always consider HCW as non-infectious
track_diag <- dplyr::filter(track_1025, is_infectious)
track_undiag <- dplyr::filter(track_undiag, !is_infectious | is.na(is_infectious))
track_tb_all <- rbind(
  track_diag,
  filter(track_undiag, patient_id %in% param_1025[[paste0("undiagTB","_", "du")]][[sim]])
)
aer_df <- readRDS("../data-clean/environmental/air-change-rate.rds") %>%
  filter(month(date) == as.numeric(mth),
         day(date) == as.numeric(dy))

# time
date <- paste0("2021-",mth, "-", dy)
t_start <- as.POSIXct(paste(date, "08:00:00"))
midday <- as.POSIXct(paste(date, "12:00:00"))
t_end <- as.POSIXct(paste(date, "16:00:00"))
tt <- seq.POSIXt(t_start, t_end, by = "1 sec")
tt_morn <- seq.POSIXt(t_start, midday, by = "1 sec")
tt_noon <- seq.POSIXt(midday + 1, t_end, by = "1 sec")
midday_idx <- which(tt == midday)

# simulation
ct_by_room <- list()
for (room in c("Waiting room", "Corridor", "TB room")) {

    print(paste0("- ", room))

    # room setup
    if (room == "Waiting room") {

      # initial quanta concentration matrix
      c0 <- sP_to_matrix(waiting_room)

@@

      # room dimension
      roomDim <- dimWR

    } else if (room == "Corridor") {

      # initial quanta concentration matrix
      c0 <- sP_to_matrix(corridor)

@@

      # room dimension
      roomDim <- dimCD

    } else if (room == "TB room") {

      # initial quanta concentration matrix
      c0 <- sP_to_matrix(tb_room)

@@

      # room dimension
      roomDim <- dimTB

    } else {
      stop("Error: Room invalid.")
    }

    # add quanta generation rate per second by activity level
    track_tb <- track_tb %>%
      dplyr::select(time, cell_x, cell_y, activity) %>%
      rename(x = cell_x, y = cell_y) %>%
      mutate(q = ifelse(activity == 1, param_1025$q_wait[param_1025$sim == sim], param_1025$q_walk[param_1025$sim == sim]),
             q = q * param_1025$mask_red[param_1025$sim == sim],
             q = q / 3600)

    # quanta conc in the morning
    print("-- morning")
    track_tb_morn <- filter(track_tb, time <= midday)
    track_tb_morn$t <- as.numeric(difftime(track_tb_morn$time, t_start, units = "sec"))

    ct_by_room[[room]] <- stm(time = tt_morn,
                     c0 = c0, 
                     inf = dplyr::select(track_tb_morn, -time, -activity),
                     cellLength = convert_dist(cellSize),
                     pd = 0,
                     vol = prod(roomDim),
                     aer = aer_df$aer[aer_df$room==room & aer_df$daytime=="Morning"] / 3600,
                     lambda = param_1025$viral_inact[param_1025$sim == sim] / 3600,
                     deposit = 0)
}

# get subset
subset_qc <- function(t) {

  # quanta conc
  qc_wr <- melt(ct_by_room[["Waiting room"]][,,t]) %>%
    rename(y = Var1, x = Var2) %>%
    mutate(room = "Waiting room", across(c(x, y), as.numeric))
  qc_cr <- melt(ct_by_room[["Corridor"]][,,t]) %>%
    rename(y = Var1, x = Var2) %>%
    mutate(room = "Corridor", across(c(x, y), as.numeric))
  qc_tb <- melt(ct_by_room[["TB room"]][,,t]) %>%
    rename(y = Var1, x = Var2) %>%
    mutate(room = "TB room", across(c(x, y), as.numeric))

  # link to building
  qc <- rbind(qc_wr, qc_cr, qc_tb) 
  qc <- left_join(roomCoord, qc, by = c("x", "y", "room")) %>%
    mutate(id = paste0("ID", id),
           t = t)

@@

track_tb_all %>%
  filter(between(time, as.POSIXct(paste(date, "09:41:00")), as.POSIXct(paste(date, "09:45:00")))) %>%
  ggplot(aes(x = x, y = y, color = factor(patient_id))) +
  geom_path()

tt_selected <- seq.POSIXt(as.POSIXct(paste(date, "09:41:00")), as.POSIXct(paste(date, "09:45:00")), by = "1 sec")
tt_selected_idx <- as.numeric(tt_selected - t_start, unit = "secs")
names(tt_selected) <- tt_selected_idx

qc_sub <- do.call(rbind, parallel::mclapply(tt_selected_idx, subset_qc, mc.cores = 4)) %>%
  mutate(t = t - min(tt_selected_idx))

# animated plot
qc_anim_pl <- ggplot(data = left_join(fortify(waiting_room), filter(qc_sub, room == "Waiting room"), by = c("id")),
         mapping = aes(x = long, y = lat, group = group, fill = value)) +
  geom_polygon() +
  coord_cartesian(expand = F) +
  scale_fill_gradientn(colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"), labels = function(x) round(1e3 * x / (0.25^2 * 3), 1)) +
  theme_custom(14) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.position = "bottom",
        panel.background = element_rect(fill = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd")[1]), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  transition_time(t) +
  labs(subtitle = "Time (secs): {frame_time}", title = "Quanta concentration in the waiting room on Oct 25 between 9:41 and 09:45am",
       fill = expression("Quanta/m"^3*" x 10e"^-3))

animate(qc_anim_pl, duration = 30, width = 1.850877 * 600, height = 600, renderer = gifski_renderer(), nframes = length(tt_selected_idx))
anim_save("../results/modeling/illustrated-quanta-conc-over-time.gif")
```

@@

```{r}
main_files <- c(
  list.files("../simulations/main/10-13/", full.names = T),
  list.files("../simulations/main/10-15/", full.names = T),
  list.files("../simulations/main/10-25/", full.names = T),
  list.files("../simulations/main/11-04/", full.names = T),
  list.files("../simulations/main/11-05/", full.names = T))
```

@@

mean_qc <- function(files, room, daytime) {
  room_files <- files[grepl(room, files)]
  room_dt_files <- room_files[grepl(daytime, room_files)]
  qc <- simplify2array(lapply(room_dt_files, readRDS))
  mqc <- apply(qc, c(1,2), mean)
  mqc_df <- melt(mqc) %>%
    rename(y = Var1,
           x = Var2) %>%
    mutate(room = room,
           daytime = daytime,
           across(c(x, y), as.numeric))
}

@@

# link to building
main_qc <- rbind(main_qc_wr_mo, main_qc_wr_af, 
                 main_qc_cr_mo, main_qc_cr_af,
                 main_qc_tb_mo, main_qc_tb_af) 
main_qc <- left_join(roomCoord, 
                     main_qc, 
                     by = c("x", "y", "room")) %>%
  mutate(id = paste0("ID", id),
         daytime = factor(ifelse(daytime == "morning", "Morning", "Afternoon"), levels = c("Morning", "Afternoon")))

# plot

mean_quanta_pl <- plot_spatial(
  ggplot(mapping = aes(x = long, y = lat, group = group, fill = value)) +
    facet_wrap(~ daytime) +
    geom_polygon(data = left_join(fortify(waiting_room), filter(main_qc, room == "Waiting room"), by = c("id"))) +
    geom_polygon(data = left_join(fortify(corridor), filter(main_qc, room == "Corridor"), by = c("id"))) +
    geom_polygon(data = left_join(fortify(tb_room), filter(main_qc, room == "TB room"), by = c("id"))) +
    labs(fill = expression("Quanta/m"^3*" x 10e"^-3), title = "Spatial quanta concentration by daytime", subtitle = "Mean concentration in the morning and afternoon") +
    scale_fill_gradientn(colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"), labels = function(x) round(1e3 * x / (0.25^2 * 3), 1)),
  14
  ) +
  theme(legend.position = c(0.915,0.095), legend.direction = "horizontal",
        legend.key.height = unit(0.2, "cm"), legend.key.width = unit(.8, "cm")) +
  guides(fill = guide_colorbar(title.position = "top"))

save_plot(mean_quanta_pl, pdf_file = "../results/modeling/quanta-concentration.pdf", w = 26, h = 12)
```

@@

```{r}
# files
main_tr_files <- main_files[grepl("risk", main_files)]

# data
main_tr <- do.call(rbind, lapply(main_tr_files, function(f) readRDS(f) %>% mutate(mthdy = stringi::stri_extract(f, regex = "\\d{2}-\\d{2}"), 
                                                                                  sim = stringi::stri_extract(f, regex = "\\d{1,6}"))))

mean_main_tr <- main_tr %>%
  mutate(patient_id = paste(mthdy, patient_id, sep = "-")) %>%
  group_by(patient_id) %>%
  summarize(P_mean = mean(P),
            P_med = median(P)) %>%
  ungroup() %>% 
  reshape2::melt("patient_id") %>%
  mutate(val_col = ifelse(value < .05, "Low", ifelse(value < .1, "Median", "High")),
         val_col = factor(val_col, levels = c("Low", "Median", "High")),
         variable = ifelse(grepl("mean", variable), "Mean risk", "Median risk"),
         variable = factor(variable, levels = c("Mean risk", "Median risk")))

roi_annon <- mean_main_tr %>%
  group_by(variable) %>%
  summarize(mean_P = mean(value),
            lower_P = quantile(value, .025),
            upper_P = quantile(value, .975)) %>%
  ungroup() %>%
  mutate(across(c(mean_P, lower_P, upper_P), ~ round_k(100 * .x, 1)),
         lab = paste0("Mean ", mean_P, "%", " (95%-CrI ", lower_P, "% – ", upper_P, "%)"),
         value = .25, 
         y = 1350)

roi_pl <- ggplot(mean_main_tr, aes(x = value)) +
  geom_histogram(mapping = aes(fill = val_col), breaks = seq(.0, .5, .05)) +
  geom_text(data = roi_annon, mapping = aes(x = value, y = y, label = lab), size = 12 / cm(1)) +
  facet_wrap(~ variable) +
  scale_fill_manual(values = wes_palette("Darjeeling1")[c(2,3,1)]) +
  coord_cartesian(expand = F) +
  scale_x_continuous(breaks = c(0, .05, 0.1, seq(0.2, .5, .1)), labels = function(x) x * 100) +
  scale_y_continuous(labels = scales::comma, breaks = seq(0, 1500, 250)) +
  labs(x = "Risk of infection (%)", y = "Number of patients",
       title = "Risk of infection",
       subtitle = "Mean and median individual risk of infection in clinic") +
  theme_custom(14) +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        panel.spacing = unit(2, "cm"))

save_plot(roi_pl, pdf_file = "../results/modeling/risk-of-infection.pdf", w = 26, h = 12)
```

@@

```{r}
# load tracking and environmental data
track <- readRDS("../data-clean/patient-tracking/augmented-and-combined-data.rds")
aer <- readRDS("../data-clean/environmental/air-change-rate.rds") %>%
  mutate(daytime = factor(daytime, levels = c("Morning", "Afternoon")))

# process tracking data
track_daytime <- track %>%
  mutate(h = lubridate::hour(time)) %>%
  group_by(date, patient_id) %>%
  summarize(dur_morning = sum(h < 12),
            dur_afternoon = sum(h >= 12),
            tracking_end = tracking_end[1]) %>%
  ungroup() %>%
  mutate(daytime = ifelse(dur_morning > dur_afternoon, "Morning", "Afternoon"))

time_in_clin <- track %>%
  group_by(date, patient_id) %>%
  summarize(`Clinic` = n(),
            `Waiting room` = sum(is_waitingroom),
            `TB room` = sum(is_tbroom),
            `Corridor` = sum(is_passage)) %>%
  ungroup() %>%
  mutate(across(c(`Clinic`, `Waiting room`, `TB room`, `Corridor`), ~ .x / 60))

@@

time_close_contacts <- track %>%
  mutate(has_closeContact = ifelse(sapply(closeContacts, length) >= 1, T, F)) %>%
  group_by(date, patient_id) %>%
  summarize(cc_time = sum(has_closeContact) / 60) %>%
  ungroup()

# merge data
track_ft <- left_join(track_daytime, left_join(left_join(time_in_clin, no_close_contacts), time_close_contacts)) %>%
  mutate(n_cc = ifelse(is.na(n_cc), 0, n_cc)) %>%
  rename(time_in_clinic = `Clinic`) %>%
  dplyr::select(date, patient_id, tracking_end, daytime, time_in_clinic, n_cc, cc_time)
assoc_df <- track_ft %>%
  left_join(aer %>% filter(room == "Waiting room") %>% dplyr::select(date, daytime, aer)) %>%
  mutate(patient_id = paste(lubridate::month(date), lubridate::day(date), patient_id, sep = "-")) %>%
  left_join(reshape2::dcast(mean_main_tr, patient_id ~ variable) %>% dplyr::select(patient_id, `Mean risk`) %>% rename(mean_risk = `Mean risk`))

# model
amod <- stan_betareg(mean_risk ~ aer + time_in_clinic + n_cc + cc_time, 
                     data = assoc_df %>% mutate(n_cc = n_cc + 1, cc_time = cc_time + 1/60,
                                                across(c(aer, time_in_clinic, n_cc, cc_time), log2)), 
                     seed = 12345, cores = 4)

@@

amod_est_pl <- ggplot(amod_est, aes(y = variable, x = value)) +
  geom_vline(aes(xintercept = 1), linetype = "dotted", color = "red") +
  stat_pointinterval() +
  labs(x = "Odds ratio associated with a doubling of the predictor",
       title = "Predictors of infection risk",
       subtitle = "Association of close contact, time in clinic, and air change rate with mean risk of infection",
       caption = "Note: Median as dots, 66%- and 95%-CrI as thick and thin lines.") +
  scale_y_discrete(labels = c("Air change rate", "Time in clinic", "Number of close contacts", "Time in clocse contact")) +
  theme_custom(14) +
  theme(axis.title.y = element_blank())

save_plot(amod_est_pl, pdf_file = "../results/modeling/association-predictors.pdf", w = 20, h = 12)
```

>>>>

source("utils/plotting.r")
source("utils/tex.r")
source("models/stm-v2.R")

# building
cellSize <- 250
source("preprocessing/prep_building-rasterize.R")

# load tracking and environmental data
track <- readRDS("data-clean/patient-tracking/augmented-and-combined-data.rds")
aer <- readRDS("data-clean/environmental/air-change-rate.rds") %>%
  mutate(daytime = factor(daytime, levels = c("Morning", "Afternoon")))
```

@@

```{r}
# data
mth <- 10
dy <- 25
sim <- 1
param_1025 <- readRDS(
  paste0(
    "../simulations/", "/", "sim-param-",
    2021, "-", mth, "-", dy, ".rds"
  )
)
track_1025 <- readRDS(
  paste0(
    "../simulations/", "/", "tracking-df-",
    250, "px-", 2021, "-", mth, "-", dy, ".rds"
  )
)
track_undiag <- track_1025 %>%
  dplyr::filter(tracking_end != "Possible HCW")
track_diag <- dplyr::filter(track_1025, is_infectious)
track_undiag <- dplyr::filter(
  track_undiag, !is_infectious | is.na(is_infectious)
)
track_tb_all <- rbind(
  track_diag,
  filter(
    track_undiag,
    patient_id %in% param_1025[[paste0("undiagTB", "_", "du")]][[sim]]
  )
)
aer_df <- readRDS("../data-clean/environmental/air-change-rate.rds") %>%
  filter(
    month(date) == as.numeric(mth),
    day(date) == as.numeric(dy)
  )

# time
date <- paste0("2021-", mth, "-", dy)
t_start <- as.POSIXct(paste(date, "08:00:00"))
midday <- as.POSIXct(paste(date, "12:00:00"))
t_end <- as.POSIXct(paste(date, "16:00:00"))
tt <- seq.POSIXt(t_start, t_end, by = "1 sec")
tt_morn <- seq.POSIXt(t_start, midday, by = "1 sec")
tt_noon <- seq.POSIXt(midday + 1, t_end, by = "1 sec")
midday_idx <- which(tt == midday)

# simulation
ct_by_room <- list()
for (room in c("Waiting room", "Corridor", "TB room")) {
  print(paste0("- ", room))

  # room setup
  if (room == "Waiting room") {
    # initial quanta concentration matrix
    c0 <- sP_to_matrix(waiting_room)

@@

    # room dimension
    roomDim <- dimWR
  } else if (room == "Corridor") {
    # initial quanta concentration matrix
    c0 <- sP_to_matrix(corridor)

@@

    # room dimension
    roomDim <- dimCD
  } else if (room == "TB room") {
    # initial quanta concentration matrix
    c0 <- sP_to_matrix(tb_room)

@@

    # room dimension
    roomDim <- dimTB
  } else {
    stop("Error: Room invalid.")
  }

  # add quanta generation rate per second by activity level
  track_tb <- track_tb %>%
    dplyr::select(time, cell_x, cell_y, activity) %>%
    rename(x = cell_x, y = cell_y) %>%
    mutate(
      q = ifelse(activity == 1,
        param_1025$q_wait[param_1025$sim == sim],
        param_1025$q_walk[param_1025$sim == sim]
      ),
      q = q * param_1025$mask_red[param_1025$sim == sim],
      q = q / 3600
    )

  # quanta conc in the morning
  print("-- morning")
  track_tb_morn <- filter(track_tb, time <= midday)
  track_tb_morn$t <- as.numeric(difftime(
    track_tb_morn$time, t_start,
    units = "sec"
  ))

  ct_by_room[[room]] <- stm(
    time = tt_morn,
    c0 = c0,
    inf = dplyr::select(track_tb_morn, -time, -activity),
    cellLength = convert_dist(cellSize),
    pd = 0,
    vol = prod(roomDim),
    aer = aer_df$aer[aer_df$room == room & aer_df$daytime == "Morning"] / 3600,
    lambda = param_1025$viral_inact[param_1025$sim == sim] / 3600,
    deposit = 0
  )
}

# get subset
subset_qc <- function(t) {
  # quanta conc
  qc_wr <- melt(ct_by_room[["Waiting room"]][, , t]) %>%
    rename(y = Var1, x = Var2) %>%
    mutate(room = "Waiting room", across(c(x, y), as.numeric))
  qc_cr <- melt(ct_by_room[["Corridor"]][, , t]) %>%
    rename(y = Var1, x = Var2) %>%
    mutate(room = "Corridor", across(c(x, y), as.numeric))
  qc_tb <- melt(ct_by_room[["TB room"]][, , t]) %>%
    rename(y = Var1, x = Var2) %>%
    mutate(room = "TB room", across(c(x, y), as.numeric))

  # link to building
  qc <- rbind(qc_wr, qc_cr, qc_tb)
  qc <- left_join(roomCoord, qc, by = c("x", "y", "room")) %>%
    mutate(
      id = paste0("ID", id),
      t = t
    )

@@

track_tb_all %>%
  filter(between(
    time,
    as.POSIXct(paste(date, "09:41:00")),
    as.POSIXct(paste(date, "09:45:00"))
  )) %>%
  ggplot(aes(x = x, y = y, color = factor(patient_id))) +
  geom_path()

tt_selected <- seq.POSIXt(
  as.POSIXct(paste(date, "09:41:00")),
  as.POSIXct(paste(date, "09:45:00")),
  by = "1 sec"
)
tt_selected_idx <- as.numeric(tt_selected - t_start, unit = "secs")
names(tt_selected) <- tt_selected_idx

qc_sub <- do.call(
  rbind,
  parallel::mclapply(tt_selected_idx, subset_qc, mc.cores = 4)
) %>%
  mutate(t = t - min(tt_selected_idx))

# animated plot
qc_anim_pl <- ggplot(
  data = left_join(
    fortify(waiting_room),
    filter(qc_sub, room == "Waiting room"),
    by = c("id")
  ),
  mapping = aes(x = long, y = lat, group = group, fill = value)
) +
  geom_polygon() +
  coord_cartesian(expand = FALSE) +
  scale_fill_gradientn(
    colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"),
    labels = function(x) round(1e3 * x / (0.25^2 * 3), 1)
  ) +
  theme_custom(14) +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "bottom",
    panel.background = element_rect(
      fill = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd")[1]
    ),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank()
  ) +
  transition_time(t) +
  labs(
    subtitle = "Time (secs): {frame_time}",
    title = "Quanta concentration in the waiting room on Oct 25 between 9:41 and 09:45am", # nolint
    fill = expression("Quanta/m"^3 * " x 10e"^-3)
  )

animate(
  qc_anim_pl,
  duration = 30,
  width = 1.850877 * 600,
  height = 600,
  renderer = gifski_renderer(),
  nframes = length(tt_selected_idx)
)

anim_save("results/modeling/illustrated-quanta-conc-over-time.gif")
```

@@

```{r}
main_files <- c(
  list.files("simulations/main/10-13/", full.names = T),
  list.files("simulations/main/10-15/", full.names = T),
  list.files("simulations/main/10-25/", full.names = T),
  list.files("simulations/main/11-04/", full.names = T),
  list.files("simulations/main/11-05/", full.names = T)
)
```

@@

mean_qc <- function(files, room, daytime) {
  room_files <- files[grepl(room, files)]
  room_dt_files <- room_files[grepl(daytime, room_files)]
  qc <- simplify2array(lapply(room_dt_files, readRDS))
  mqc <- apply(qc, c(1, 2), mean)
  mqc_df <- melt(mqc) %>%
    rename(
      y = Var1,
      x = Var2
    ) %>%
    mutate(
      room = room,
      daytime = daytime,
      across(c(x, y), as.numeric)
    )
}

@@

# link to building
main_qc <- rbind(
  main_qc_wr_mo, main_qc_wr_af,
  main_qc_cr_mo, main_qc_cr_af,
  main_qc_tb_mo, main_qc_tb_af
)
main_qc <- left_join(roomCoord,
  main_qc,
  by = c("x", "y", "room")
) %>%
  mutate(
    id = paste0("ID", id),
    daytime = factor(
      ifelse(daytime == "morning", "Morning", "Afternoon"),
      levels = c("Morning", "Afternoon")
    )
  )

# plot
mean_quanta_pl <- plot_spatial(
  ggplot(mapping = aes(x = long, y = lat, group = group, fill = value)) +
    facet_wrap(~daytime) +
    geom_polygon(
      data = left_join(
        fortify(waiting_room),
        filter(main_qc, room == "Waiting room"),
        by = c("id")
      )
    ) +
    geom_polygon(
      data = left_join(
        fortify(corridor),
        filter(main_qc, room == "Corridor"),
        by = c("id")
      )
    ) +
    geom_polygon(
      data = left_join(fortify(tb_room),
        filter(main_qc, room == "TB room"),
        by = c("id")
      )
    ) +
    labs(
      fill = expression("Quanta concentration in quanta/m"^3 * " x 10e"^-3),
      title = "a"
    ) +
    scale_fill_gradientn(
      colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"),
      labels = function(x) round(1e3 * x / (0.25^2 * 3), 1)
    ),
  8
) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.height = unit(0.2, "cm"),
    legend.key.width = unit(2, "cm")
  ) +
  guides(fill = guide_colorbar(title.position = "top"))

mean_quanta_pl
```

@@

```{r}
average_trans_risk <- function(files) {
  # files
  tr_files <- files[grepl("risk", files)]

  # data
  tr <- do.call(rbind, lapply(tr_files, function(f) {
    readRDS(f) %>% mutate(
      mthdy = stringi::stri_extract(f, regex = "\\d{2}-\\d{2}"),
      sim = stringi::stri_extract(f, regex = "\\d{1,6}")
    )
  }))

  average_tr <- tr %>%
    mutate(patient_id = paste(mthdy, patient_id, sep = "-")) %>%
    group_by(patient_id) %>%
    summarize(
      P_mean = mean(P),
      P_med = median(P)
    ) %>%
    ungroup() %>%
    left_join(track %>%
      mutate(
        mth = lubridate::month(time),
        dy = lubridate::day(time)
      ) %>%
      group_by(patient_id, mth, dy) %>%
      slice(1) %>%
      ungroup() %>%
      mutate(dy = ifelse(nchar(dy) == 1, paste0("0", dy), dy)) %>%
      mutate(patient_id = paste(
        mth, dy,
        patient_id,
        sep = "-"
      )) %>%
      dplyr::select(patient_id, tracking_end)) %>%
    filter(tracking_end != "Possible HCW") %>%
    dplyr::select(-tracking_end) %>%
    reshape2::melt("patient_id") %>%
    mutate(
      val_col = ifelse(value < .05, "Low",
        ifelse(value < .1, "Median", "High")
      ),
      val_col = factor(val_col, levels = c("Low", "Median", "High")),
      variable = ifelse(
        grepl("mean", variable),
        "Mean risk of infection",
        "Median risk of infection"
      ),
      variable = factor(
        variable,
        levels = paste(c("Mean", "Median"), "risk of infection")
      )
    )

  return(average_tr)
}

# average risk
main_tr <- average_trans_risk(main_files)

# annotation
annotate_average_trans_risk <- function(tr) {
  roi_annon <- tr %>%
    group_by(variable) %>%
    summarize(
      mean_P = mean(value),
      lower_P = quantile(value, .025),
      upper_P = quantile(value, .975)
    ) %>%
    ungroup() %>%
    mutate(across(c(mean_P, lower_P, upper_P), ~ round_k(100 * .x, 1)),
      lab = paste0(
        "Average ", mean_P, "%",
        " (95%-CrI ", lower_P, "% – ", upper_P, "%)"
      )
    )

  return(roi_annon)
}

main_annon <- annotate_average_trans_risk(main_tr)

# plot
roi_pl <- ggplot(main_tr, aes(x = value)) +
  geom_histogram(
    mapping = aes(fill = val_col),
    binwidth = .05,
    boundary = 1
  ) +
  geom_text(
    data = main_annon,
    mapping = aes(x = Inf, y = Inf, label = lab),
    size = 8 / cm(1),
    hjust = 1, vjust = 2
  ) +
  facet_wrap(~variable) +
  scale_fill_manual(values = wes_palette("Darjeeling1")[c(2, 3, 1)]) +
  coord_cartesian(expand = FALSE) +
  scale_x_continuous(
    breaks = c(0, .05, 0.1, seq(0.2, .5, .1)),
    labels = function(x) x * 100
  ) +
  scale_y_continuous(
    labels = scales::comma
  ) +
  labs(
    x = "Risk of infection (%)",
    y = "Number of clinical attendees",
    title = "b"
  ) +
  theme_custom() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.key.height = unit(.6, "cm"),
    legend.key.width = unit(.4, "cm"),
    panel.spacing = unit(2, "cm")
  )

roi_pl
```

@@

```{r}
# process tracking data
track_daytime <- track %>%
  mutate(h = lubridate::hour(time)) %>%
  group_by(date, patient_id) %>%
  summarize(
    dur_morning = sum(h < 12),
    dur_afternoon = sum(h >= 12),
    tracking_end = tracking_end[1]
  ) %>%
  ungroup() %>%
  mutate(daytime = ifelse(dur_morning > dur_afternoon, "Morning", "Afternoon"))

time_in_clin <- track %>%
  group_by(date, patient_id) %>%
  summarize(
    `Clinic` = n(),
    `Waiting room` = sum(is_waitingroom),
    `TB room` = sum(is_tbroom),
    `Corridor` = sum(is_passage)
  ) %>%
  ungroup() %>%
  mutate(across(c(`Clinic`, `Waiting room`, `TB room`, `Corridor`), ~ .x / 60))

@@

time_close_contacts <- track %>%
  mutate(has_closeContact = ifelse(
    sapply(closeContacts, length) >= 1, TRUE, FALSE
  )) %>%
  group_by(date, patient_id) %>%
  summarize(cc_time = sum(has_closeContact) / 60) %>%
  ungroup()

# merge data
track_ft <- left_join(
  track_daytime,
  left_join(left_join(
    time_in_clin,
    no_close_contacts
  ), time_close_contacts)
) %>%
  mutate(n_cc = ifelse(is.na(n_cc), 0, n_cc)) %>%
  rename(time_in_clinic = `Clinic`) %>%
  dplyr::select(
    date,
    patient_id,
    tracking_end,
    daytime,
    time_in_clinic,
    n_cc,
    cc_time
  )

assoc_df <- track_ft %>%
  left_join(
    aer %>%
      filter(room == "Waiting room") %>%
      dplyr::select(date, daytime, aer)
  ) %>%
  mutate(patient_id = paste(
    lubridate::month(date),
    lubridate::day(date),
    patient_id,
    sep = "-"
  )) %>%
  left_join(
    reshape2::dcast(main_tr, patient_id ~ variable) %>%
      dplyr::select(patient_id, `Mean risk of infection`) %>%
      rename(mean_risk = `Mean risk of infection`)
  )

# model
amod <- stan_betareg(mean_risk ~ aer + time_in_clinic + n_cc + cc_time,
  data = assoc_df %>% mutate(
    n_cc = n_cc + 1, cc_time = cc_time + 1 / 60,
    across(c(aer, time_in_clinic, n_cc, cc_time), log2)
  ),
  seed = 12345, cores = 4
)

@@

amod_est_pl <- ggplot(amod_est, aes(y = variable, x = value)) +
  geom_vline(aes(xintercept = 1), linetype = "dotted", color = "red") +
  stat_interval() +
  stat_summary(
    geom = "point",
    fun = median,
    size = 2,
    shape = 21,
    fill = "white"
  ) +
  labs(
    x = "Odds ratio of the modeled risk of infection for a doubling in the predictor",
    title = "c",
    color = "Credible interval"
  ) +
  scale_y_discrete(
    labels = c(
      "Air change rate",
      "Time in clinic",
      "Number of close contacts",
      "Time in clocse contact"
    )
  ) +
  scale_color_brewer(
    labels = paste0(c(95, 80, 50), "%")
  ) +
  theme_custom() +
  theme(axis.title.y = element_blank())

amod_est_pl
```

```{r main-figure}
main_fig <- grid.arrange(
  mean_quanta_pl,
  roi_pl,
  amod_est_pl,
  ncol = 1,
  heights = c(8, 8, 4)
)

ggsave(
  main_fig,
  filename = "results/modeling/main-figure.png",
  width = 16 / cm(1),
  height = 20 / cm(1)
)
```

## Additional analyses

```{r additional-analyses}
# files
files_nm <- list.files("simulations/nomasks/",
  full.names = TRUE, pattern = "*.rds$",
  recursive = TRUE
)
files_vent <- list.files("simulations/ventilation/",
  full.names = TRUE, pattern = "*.rds$",
  recursive = TRUE
)
files_temp <- list.files("simulations/temporal/",
  full.names = TRUE, pattern = "*.rds$",
  recursive = TRUE
)
files_prev <- list.files("simulations/prevalence/",
  full.names = TRUE, pattern = "*.rds$",
  recursive = TRUE
)
files_susp <- list.files("simulations/suspected/",
  full.names = TRUE, pattern = "*.rds$",
  recursive = TRUE
)

# average risk
nm_tr <- average_trans_risk(files_nm)
vent_tr <- average_trans_risk(files_vent)
temp_tr <- average_trans_risk(files_temp)
prev_tr <- average_trans_risk(files_prev)
susp_tr <- average_trans_risk(files_susp)

tr_comp <- rbind(
  nm_tr %>% mutate(scenario = "No mask wearing"),
  vent_tr %>% mutate(scenario = "Lower ventilation"),
  prev_tr %>% mutate(scenario = "Prevalence-based model"),
  susp_tr %>% mutate(scenario = "Suspected-based model")
) %>%
  mutate(
    scenario = factor(scenario, levels = c(
      "Spatiotemporal model",
      "Temporal model",
      "No mask wearing",
      "Lower ventilation",
      "Prevalence-based model",
      "Suspected-based model"
    ))
  )

# annotation
nm_annon <- annotate_average_trans_risk(nm_tr)
vent_annon <- annotate_average_trans_risk(vent_tr)
temp_annon <- annotate_average_trans_risk(temp_tr)
prev_annon <- annotate_average_trans_risk(prev_tr)
susp_annon <- annotate_average_trans_risk(susp_tr)

annon_comp <- rbind(
  nm_annon %>% mutate(scenario = "No mask wearing"),
  vent_annon %>% mutate(scenario = "Lower ventilation"),
  prev_annon %>% mutate(scenario = "Prevalence-based model"),
  susp_annon %>% mutate(scenario = "Suspected-based model")
) %>%
  mutate(
    scenario = factor(scenario, levels = c(
      "No mask wearing",
      "Lower ventilation",
      "Prevalence-based model",
      "Suspected-based model"
    ))
  )

# plot mean
mean_roi_comp_pl <- ggplot(
  tr_comp %>%
    filter(variable == "Mean risk of infection"),
  aes(x = value)
) +
  geom_histogram(
    mapping = aes(fill = val_col),
    binwidth = .05,
    boundary = 1
  ) +
  geom_text(
    data = annon_comp %>% filter(variable == "Mean risk of infection"),
    mapping = aes(x = Inf, y = Inf, label = lab),
    size = 8 / cm(1),
    hjust = 1, vjust = 2
  ) +
  facet_wrap(~scenario, ncol = 2) +
  scale_fill_manual(values = wes_palette("Darjeeling1")[c(2, 3, 1)]) +
  coord_cartesian(expand = FALSE) +
  scale_x_continuous(
    breaks = c(0, .05, 0.1, seq(0.2, .5, .1)),
    labels = function(x) x * 100
  ) +
  scale_y_continuous(
    labels = scales::comma
  ) +
  labs(
    x = "Risk of infection (%)",
    y = "Number of clinical attendees",
    title = "b"
  ) +
  theme_custom() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.key.height = unit(.6, "cm"),
    legend.key.width = unit(.4, "cm"),
    panel.spacing = unit(2, "cm")
  )

save_plot(
  mean_roi_comp_pl,
  "results/modeling/mean-roi-comparison.png",
  w = 16,
  h = 12
)

# comparison with temporal model
mod_comp <- left_join(
  temp_tr %>% rename(`Temporal model` = value) %>% dplyr::select(-val_col),
  main_tr %>% mutate(`Spatiotemporal model` = value) %>% dplyr::select(-val_col)
) %>%
  mutate(
    diff = `Spatiotemporal model` - `Temporal model`
  ) %>%
  mutate(
    val_col = ifelse(diff > .005, "Higher",
      ifelse(diff < -.005, "Lower", "Similar")
    ),
    val_col = factor(val_col, levels = c("Lower", "Similar", "Higher"))
  )
